Evolution of ORM Systems: Formal Foundations

### 2013-02-02 doc/description.tex
case study

\paragraph{Merge of Classes} There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used. It can be used in case structure of both classes (respectively tables) are equal. The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
\begin{align*}
&	\mathbf{Merge} : Label \times Label \times Label \times State \rightarrow State \\
&	\Phi(s, Merge(l_1, l_2, l_n) = \\
& \;\;\;\;\;\; = \inference{
		\exists c_1, c_2 \in classes(application(s)) \wedge 
		equals(c_1, c_2)}{\begin{gathered}
			state(application(s)) \\ \circ addClass(cl(ln, properties(c_1), associations(c_1)) \\
			\circ (\forall c \in classes (application(s)), a \in associations(c) \\
			reference(a) = c_1 \vee reference(a) = c_2 : \\
			reference(a) = l_n) \\
			\circ renClass(l_1) \circ renClass(l_2) \end{gathered}}
 \\\\
&	\Psi(s, merge(l_1, l_2, l_n)) = \\
& \;\;\;\;\;\; = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)\\
\forall f \in (foreignKeys(t_1) \cup foreignKey(t_2)) \\ \setminus (foreignKey(t_1) \cap foreignKey(t_2)) \\ : NOTNULL \not \in constraints(f)  }{\begin{gathered}
	state(database(s)) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
\bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 
\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(select(name(f), k)) = pkValue(r) : \\value(select(name(f), k)) = pkValue(n))) \\  \bullet dropTableAndData(l_1) \bullet dropTableAndData(l_2)
 \end{gathered}} \\\\
 & l_1, l_2, l_n \in Label, s \in State
\end{align*}

\section{Case Study}
The case study demonstrates how the transformation helps with data evolution in real example of software evolution. There are two classes in our software $Person$ and $LegalPerson$ both of them contains information about address (street, city and zip code). The situation in software is in Figure \ref{fig:case1}, where the whole software is shown.

@@

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{./images/case_app_1}
	\caption{The application model.}
\end{subfigure}
\quad
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{./images/case_db_1}
	\caption{The database model.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l | }
	 	\hline
		Id &  Surname & Street & City & ZIP  \\ \hline  
		11 & Jackson & Central Park St & New York & 100 01  \\ \hline
		12 & Clooney & S Orange Ave & Orlando & 320 24  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Person$.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l | c |}
	 	\hline
		Id &  BusinessName & Street & City & ZIP \\ \hline  
		100 & Tools \& Machines & Olive ave. & New York & 100 01 \\ \hline
		200 & AI Robotics & Pine ave. & LA & 900 03  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $LegalPerson$.}
\end{subfigure}
	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalPerson$ class.}
	\label{fig:case1}
\end{figure}

The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalPerson$. Temporary classes has to be connected with the origin classes by an association. The result is shown in Figure \ref{fig:case2}.
\begin{align*}
S_1 & = state(S) \\
& \;\;\;\;\;\; \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") \\
& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) \\
&\;\;\;\;\;\; \oplus Extract("LegalPerson", "Address\_tmp2", "street", \\
& \;\;\;\;\;\;\;\;\;\;\;\; "city", "zip")  \\
& \;\;\;\;\;\; \oplus AddAssociation("LegalPerson", \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; assoc("residentialAddress", "Address\_tmp2", 1, 1),\\ 
& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("LegalPerson", database(S)), \\
& \;\;\;\;\;\;\;\;\;\;\;\; y \in instances("Address\_tmp2", database(S)) : \\
& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) \\
\end{align*}

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{./images/case_app_5}
	\caption{The application model.}
\end{subfigure}
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{./images/case_db_5}
	\caption{The database model.}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
	\centering
	\begin{tabular}{| c | l |}
	 	\hline
		Id &  Surname \\ \hline  
		11 & Jackson  \\ \hline
		12 & Clooney  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Person$.}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
	\centering
	\begin{tabular}{| c | l |}
	 	\hline
		Id &  BusinessName  \\ \hline  
		100 & Tools \& Machines  \\ \hline
		200 & AI Robotics \\ \hline
	\end{tabular}
	\caption{Data stored in the table $LegalPerson$.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l |}
	 	\hline
		Id & Street & City & ZIP & Person \\ \hline  
		11 & Central Park St & New York & 100 01 & 11  \\ \hline
		12 & Orange Ave & Orlando & 320 24 & 12 \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Address\_tmp1$.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l |}
	 	\hline
		Id & Street & City & ZIP & LegalPerson\\ \hline  
		100 & Olive ave. & New York & 100 01 & 100 \\ \hline
		200 & Pine ave. & LA & 900 03 & 200  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Address\_tmp2$.}
\end{subfigure}

	\caption{The state of the software after the information are extracted into separated classes and associations are created.}
	\label{fig:case2}
\end{figure}
%Next two evolution steps are bit intuitive but necessary for the next evolution.  then the temporary class can be merged together and finally the temporary class has to be renamed to $Address$:
The two temporary classes can be merged together as they have the same structure. All associations pointing to one of the temporary classes are redirected to the newly created class.
\begin{align*}
%
%& \oplus AddAssociation("Person", \\
%& assoc("address", "Address\_tmp2", 1, 1), \varnothing) \\
%& \oplus AddAssociation("LegalPerson", \\
%& assoc("residentialAddress", "Address\_tmp1", 1, 1),\varnothing) \\
%
S_2 &= S_1 \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
\end{align*}
Finally the associations are renamed and the evolution end.
\begin{align*}
S_3 &= S_2  \oplus RenameAssociation("Person", "Address\_tmp1", "Address")\\
& \oplus RenameAssociation("LegalPerson", "Address\_tmp2". "Address")\\
\end{align*}

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./images/case_app_9}
	\caption{The application model.}
\end{subfigure}
\quad
\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./images/case_db_8}
	\caption{The database model.}
\end{subfigure}

\begin{subfigure}[b]{0.5\textwidth}
	\centering
	\begin{tabular}{| c | l |}
	 	\hline
		Id &  Surname \\ \hline  
		11 & Jackson  \\ \hline
		12 & Clooney  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Person$.}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
	\centering
	\begin{tabular}{| c | l |}
	 	\hline
		Id &  BusinessName  \\ \hline  
		100 & Tools \& Machines  \\ \hline
		200 & AI Robotics \\ \hline
	\end{tabular}
	\caption{Data stored in the table $LegalPerson$.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l | l |}
	 	\hline
		Id & Street & City & ZIP & Person & LegalPerson\\ \hline  
		1 & Olive ave. & New York & 100 01 & & 100 \\ \hline
		2 & Pine ave. & LA & 900 03 & & 200  \\ \hline
		3 & Central Park St & New York & 100 01 & 11 &  \\ \hline
		4 & Orange Ave & Orlando & 320 24 & 12 &\\ \hline
	\end{tabular}
	\caption{Data stored in the table $Address\_tmp2$.}
\end{subfigure}

	\caption{The final state of the case study example. There is one class containing information about address for the whole software and data were migrated.}
\end{figure}

The scenario was completed successfully, because there were no constraints defined in the database and the data was suitable for the evolution there were no complication. The example of evolution presented in case study shows the possibilities of defined transformation. The example itself is not large, nevertheless it shows advanced refactoring cases, which are very often used in real software development. We believe that composition of proposed transformation can be used even in case of large data evolution cases. It is obvious that usage of proposed set of transformations only will cause very large scripts describing data evolution, therefore a mechanism for creating templates (reusable evolution scripts).

@@

\newpage
\appendix
\section{Assigment for NII Internship}
The model of evolution in ORM software defined in this paper will be improved and validated during internship at NII. The main objectives of the internship are:
\begin{enumerate}
	\item \label{it:1} Verify the model by an implementation in the Maude tool. \emph{(Estimated time: 1 month)}
	\item \label{it:2} More complex examples of refactoring cases should be examined with existing model.  \emph{(Estimated time: 0.5 month)}
	\item \label{it:3} The inheritance relationship between classes should be implemented in the model together with inheritance-based refactoring cases. \emph{(Estimated time: 0.5 month)}
	\item Prepare an article, which will contain topics \ref{it:1}, \ref{it:2} and \ref{it:3}. \emph{(Estimated time: 1 month)}
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Case Study}
\label{sec:caseStudy}
Initial State
\hline
\begin{align*}
&	S = sw(a, db(t,r))
\end{align*}
\begin{align*}
a &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing))
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing)) 
\end{align*}
\begin{align*}
r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
& row("legalPerson", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
\end{align*}
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{align*}
& state(S) \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip") = \\
&=
\end{align*}
\hline
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalPerson, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),\\
& cl("Address\_tmp1", \varnothing, \varnothing)
 )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalPerson, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab("Address\_tmp1", pk("Id"), \varnothing, \varnothing)) 
\end{align*}
\begin{align*}
r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
& row("legalPerson", ["Id", 10], \\
& ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("legalPerson", ["Id", 20], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
\end{align*}

>>>>

\paragraph{Merge of Classes} There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used. It can be used in case structure of both classes (respectively tables) are equal. The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
\begin{align*}
&	\mathbf{Merge} : Label \times Label \times Label \times State \rightarrow State \\
&	\Phi(s, Merge(l_1, l_2, l_n) = \\
& \;\;\;\;\;\; = \inference{
		\exists c_1, c_2 \in classes(application(s)) \wedge 
		equals(c_1, c_2)}{\begin{gathered}
			state(application(s)) \\ \circ addClass(cl(ln, properties(c_1), associations(c_1)) \\
			\circ (\forall c \in classes (application(s)), a \in associations(c) \\
			reference(a) = c_1 \vee reference(a) = c_2 : \\
			reference(a) = l_n) \\
			\circ remClass(l_1) \circ remClass(l_2) \end{gathered}}
 \\\\
&	\Psi(s, merge(l_1, l_2, l_n)) = \\
& \;\;\;\;\;\; = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)\\
\forall f \in (foreignKeys(t_1) \cup foreignKey(t_2)) \\ \setminus (foreignKey(t_1) \cap foreignKey(t_2)) \\ : NOTNULL \not \in constraints(f)  }{\begin{gathered}
	state(database(s)) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
\bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 
\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(select(name(f), k)) = pkValue(r) : \\value(select(name(f), k)) = pkValue(n))) \\  \bullet dropTableAndData(l_1) \bullet dropTableAndData(l_2)
 \end{gathered}} \\\\
 & l_1, l_2, l_n \in Label, s \in State
\end{align*}

\section{Case Study}
The case study demonstrates how the transformation helps with data evolution in real example of software evolution. There are two classes in our software $Person$ and $LegalParty$ both of them contains information about address (street, city and zip code). The situation in software is in Figure \ref{fig:case1}, where the whole software is shown.

@@

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{./images/case_app_1}
	\caption{The application model.}
\end{subfigure}
\quad
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{./images/case_db_1}
	\caption{The database model.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l | }
	 	\hline
		Id &  Surname & Street & City & ZIP  \\ \hline  
		11 & Jackson & Central Park St & New York & 100 01  \\ \hline
		12 & Clooney & S Orange Ave & Orlando & 320 24  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Person$.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l | c |}
	 	\hline
		Id &  BusinessName & Street & City & ZIP \\ \hline  
		100 & Tools \& Machines & Olive ave. & New York & 100 01 \\ \hline
		200 & AI Robotics & Pine ave. & LA & 900 03  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $LegalParty$.}
\end{subfigure}
	\caption{The initial state of the example used in the case study. There is a repetition of information structure in the $Person$ and $LegalParty$ class.}
	\label{fig:case1}
\end{figure}

The first  step of is to extract two temporary classes representing address of a $Person$ or of a $LegalParty$. Temporary classes has to be connected with the origin classes by an association. The result is shown in Figure \ref{fig:case2}.
\begin{align*}
S_1 & = state(S) \\
& \;\;\;\;\;\; \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip",\\ 
& \;\;\;\;\;\; \;\;\;\;\;\;  \{(x,x) | x \in instances("Person", database(S))\}) \\
& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) \\
&\;\;\;\;\;\; \oplus Extract("LegalParty", "Address\_tmp2", "street", \\
& \;\;\;\;\;\;\;\;\;\;\;\; "city", "zip")  \\
& \;\;\;\;\;\; \oplus AddAssociation("LegalParty", \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp2", 1, 1),\\ 
& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("LegalParty", database(S)), \\
& \;\;\;\;\;\;\;\;\;\;\;\; y \in instances("Address\_tmp2", database(S)) : \\
& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) \\
\end{align*}

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{./images/case_app_5}
	\caption{The application model.}
\end{subfigure}
\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{./images/case_db_5}
	\caption{The database model.}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
	\centering
	\begin{tabular}{| c | l |}
	 	\hline
		Id &  Surname \\ \hline  
		11 & Jackson  \\ \hline
		12 & Clooney  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Person$.}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
	\centering
	\begin{tabular}{| c | l |}
	 	\hline
		Id &  BusinessName  \\ \hline  
		100 & Tools \& Machines  \\ \hline
		200 & AI Robotics \\ \hline
	\end{tabular}
	\caption{Data stored in the table $LegalParty$.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l |}
	 	\hline
		Id & Street & City & ZIP & Person\_Address \\ \hline  
		11 & Central Park St & New York & 100 01 & 11  \\ \hline
		12 & Orange Ave & Orlando & 320 24 & 12 \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Address\_tmp1$.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l |}
	 	\hline
		Id & Street & City & ZIP & LegalParty\_Address\\ \hline  
		100 & Olive ave. & New York & 100 01 & 100 \\ \hline
		200 & Pine ave. & LA & 900 03 & 200  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Address\_tmp2$.}
\end{subfigure}

	\caption{The state of the software after the information are extracted into separated classes and associations are created.}
	\label{fig:case2}
\end{figure}
%Next two evolution steps are bit intuitive but necessary for the next evolution.  then the temporary class can be merged together and finally the temporary class has to be renamed to $Address$:
The two temporary classes can be merged together as they have the same structure. All associations pointing to one of the temporary classes are redirected to the newly created class.
\begin{align*}
%
%& \oplus AddAssociation("Person", \\
%& assoc("address", "Address\_tmp2", 1, 1), \varnothing) \\
%& \oplus AddAssociation("LegalParty", \\
%& assoc("residentialAddress", "Address\_tmp1", 1, 1),\varnothing) \\
%
S_2 &= S_1 \oplus Merge("Address\_tmp1", "Address\_tmp2", "Address") \\
\end{align*}
\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./images/case_app_9}
	\caption{The application model.}
\end{subfigure}
\quad
\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./images/case_db_8}
	\caption{The database model.}
\end{subfigure}

\begin{subfigure}[b]{0.5\textwidth}
	\centering
	\begin{tabular}{| c | l |}
	 	\hline
		Id &  Surname \\ \hline  
		11 & Jackson  \\ \hline
		12 & Clooney  \\ \hline
	\end{tabular}
	\caption{Data stored in the table $Person$.}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
	\centering
	\begin{tabular}{| c | l |}
	 	\hline
		Id &  BusinessName  \\ \hline  
		100 & Tools \& Machines  \\ \hline
		200 & AI Robotics \\ \hline
	\end{tabular}
	\caption{Data stored in the table $LegalParty$.}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
	\centering
	\begin{tabular}{| c | l | l | l | l | l |}
	 	\hline
		Id & Street & City & ZIP & Person\_Address & LegalParty\_Address\\ \hline  
		1 & Olive ave. & New York & 100 01 & & 100 \\ \hline
		2 & Pine ave. & LA & 900 03 & & 200  \\ \hline
		3 & Central Park St & New York & 100 01 & 11 &  \\ \hline
		4 & Orange Ave & Orlando & 320 24 & 12 &\\ \hline
	\end{tabular}
	\caption{Data stored in the table $Address\_tmp2$.}
\end{subfigure}

	\caption{The final state of the case study example. There is one class containing information about address for the whole software and data were migrated.}
\end{figure}
The scenario was completed successfully, because there were no constraints defined in the database and the data was suitable for the evolution there were no complication. The example of evolution presented in case study shows the possibilities of defined transformation. The example itself is not large, nevertheless it shows advanced refactoring cases, which are very often used in real software development. We believe that composition of proposed transformation can be used even in case of large data evolution cases. It is obvious that usage of proposed set of transformations only will cause very large scripts describing data evolution, therefore a mechanism for creating templates (reusable evolution scripts).

@@

\newpage
\appendix
\section{Assigment for NII Internship}
The model of evolution in ORM software defined in this paper will be improved and validated during internship at NII. The main objectives of the internship are:
\begin{enumerate}
	\item \label{it:1} The inheritance relationship between classes should be implemented in the model together with inheritance-based refactoring cases. \emph{(Estimated time: 0.5 month)}
	\item \label{it:2} Verify the model by an implementation in the Maude tool. \emph{(Estimated time: 1 month)}
	\item \label{it:3} More complex examples of refactoring cases should be examined with existing model.  \emph{(Estimated time: 0.5 month)}
	\item Prepare an article, which will contain topics \ref{it:1}, \ref{it:2} and \ref{it:3}. \emph{(Estimated time: 1 month)}
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Case Study}
\label{sec:caseStudy}
The first part of the evolution case study extract a new temporary class and then creates an association.
\begin{align*}
S_1 &= state(S) \\
& \;\;\;\;\;\; \oplus Extract("Person", "Address\_tmp1", "street", "city", "zip",\\ 
& \;\;\;\;\;\; \;\;\;\;\;\;  \{(x,x) | x \in instances("Person", database(S)\}) \\
& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \}) = \\
& \;\;\;\;\;\; =  state(S) \oplus addClass(cl("Address\_tmp1", \varnothing, \varnothing)) \\
& \oplus MoveProperty("Person", "Address\_tmp1", "street", "city", "zip", \\
& \;\;\;\;\;\; \;\;\;\;\;\;    \{(x,x) | x \in instances("Person", database(S))\}) \\
& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \})
\end{align*}

\newpage
Initial state of the software:
\medskip
\hline
\begin{align*}
&	S = sw(a, db(t,r))
\end{align*}
\begin{align*}
a &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing))
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing)) 
\end{align*}
\begin{align*}
r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
& row("person", ["Id", 2], ["surname", "Doe"], \\
& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
& row("LegalParty", ["Id", 10], ["businessName", "Tools and Machines ltd."], \\
& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
\end{align*}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{align*}
S_1 &= state(S) \oplus \mathbf{addClass}(cl("Address\_tmp1", \varnothing, \varnothing)) \\
& \oplus MoveProperty("Person", "Address\_tmp1", "street", "city", "zip", \\
& \;\;\;\;\;\; \;\;\;\;\;\;    \{(x,x) | x \in instances("Person", database(S))\}) \\
& \;\;\;\;\;\; \oplus AddAssociation("Person", \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; assoc("address", "Address\_tmp1", 1, 1),\\ 
& \;\;\;\;\;\;\;\;\;\;\;\; \{(x,y)| x \in instances("Person", database(S)), \\
& \;\;\;\;\;\; \;\;\;\;\;\; y \in instances("Address\_tmp1", database(S)) : \\ 
& \;\;\;\;\;\;\;\;\;\;\;\; pkVal(x) = pkVal(y) \})
\end{align*}
\hline
\begin{align*}
a' &= app(cl(Person, (pr(surname, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)), \\
& cl(LegalParty, (pr(businessName, APPSTRING, \varnothing), \\
& pr(street, APPSTRING, \varnothing), \\
& pr(city, APPSTRING, \varnothing)), pr(zip, APPSTRING, \varnothing)),\\
& \mathbf{cl("Address\_tmp1", \varnothing, \varnothing)}
 )
\end{align*}
\begin{align*}
t &= db(tab(Person, pk("Id"), col(surname, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& tab(LegalParty, pk("Id"),  col(businessName, APPSTRING, \varnothing, \varnothing), \\
& col(street, APPSTRING, \varnothing, \varnothing), col(city, APPSTRING, \varnothing, \varnothing) \\
& col(zip, APPSTRING, \varnothing, \varnothing), \varnothing), \\
& \mathbf{tab("Address\_tmp1", pk("Id"), \varnothing, \varnothing)}) 
\end{align*}
%\begin{align*}
%r &= row("person", ["Id", 1], ["surname", "Griffin"], \\
%& ["street", "S. Mary Evenue 22"], ["city", "NY"], ["zip", 10001]), \\
%& row("person", ["Id", 2], ["surname", "Doe"], \\
%& ["street", "Olive ave. 2225"], ["city", "LA"], ["zip", 90003]), \\
%& row("LegalParty", ["Id", 10], \\
%& ["businessName", "Tools and Machines ltd."], \\
%& ["street", "Hill stret 22"], ["city", "NY"], ["zip", 10001]), \\
%& row("LegalParty", ["Id", 20], ["businessName", "Doe"], \\
%& ["street", "Pine ave. 2225"], ["city", "LA"], ["zip", 90003]))
%\end{align*}
